\documentclass[11pt]{article}
\input{../suite_preamble}
\usepackage{graphicx}
\input{../suite_macros}

\title{Reflective Non-Exhaustion:\\[0.35em]
\large Certification, Realization, Fibers, and Obstruction}
\author{Nova Spivack}
\date{2026}

\begin{document}

\maketitle

\begin{abstract}
\raggedright
\textbf{Central claim.}
We state and defend a summit-level claim from a machine-checked program in Lean~4:
\emph{canonical certification does not, in general, exhaust realization}. When a
formal architecture has a bare certification layer, an enriched realization layer,
and a sound comparison map $\pi : E \to B$, collapse at the bare carrier may coexist
with nontrivial realized structure above it; the residue is organized by
non-injective comparison, nontrivial fibers, sections, and obstruction laws.

\textbf{Suite support.}
The claim is supported by six completed kinds of discharge: the internal origin
theorem in Infinity Compression; external transfer across twelve benchmark families;
an algebraic obstruction theorem for group extensions; an arithmetic bridge through
embedding problems; a topological discharge via Quillen's Theorem~A for Galois
connections; and a computability anchor (halting and Rice-style barriers in
\texttt{ComputablePred}).

\textbf{This paper's role.}
It does not introduce new axioms or replace the companion formal papers. It distills
the architecture into one summit statement, records what has been proved, and
separates the claim from incompleteness and generic proof relevance in dependent type
theory. It also explains why the collapse/refinement/fiber/obstruction pattern is a
reusable research instrument.

\textbf{Frontier.}
The next step is a strongest universal inevitability theorem over the broadest
abstract class; this manuscript is synthesis of completed instances and a candidate
general law, not that closure.

\textbf{Trust boundary:} Summit prose compresses completed formal discharges; it is not a substitute for the companion papers' Lean anchors or for future strongest-class universal theorems.
\end{abstract}

\SuiteSeriesContext{This manuscript is the \emph{summit statement}.}

\keywords{formal verification, certification, realization, fibers, obstruction,
  reflective systems, Lean~4, Mathlib}
\amsclassification{03B70, 18A30, 18G50, 68V15, 68V20}

\codeavailability{%
  \url{https://github.com/novaspivack/infinity-compression-lean};
  build \texttt{lake build};
  see \texttt{MANIFEST.md} and \texttt{lakefile.lean} for entry points and toolchain pin.}

\section{Premises, disagreement coordinates, and trust boundary}\label{sec:premises-ic-summit}
\begin{keybox}{Where disagreement can (and cannot) live}
\textbf{Formal spine.}
Completed-instance claims trace to named Lean files in the Infinity Compression program; the ``frontier'' paragraph records what is \emph{not} yet a single abstract universality theorem.

\medskip
\textbf{Premise coordinates (for external readers).}
{\raggedright
\begin{itemize}[leftmargin=2em]
\item \textbf{Instance vs.\ universal law.} The summit organizes proved cases and candidate law; it does not claim to quantify over every possible formal framework.
  \emph{Logical effect:} Opting out of the program's architecture is not a mathematical inconsistency claim.
\item \textbf{Artifact drift.} Tags and default modules may shift; use repository pins with this PDF.
  \emph{Logical effect:} Prefer the appendix theorem index in each formal paper over prose-only quotation.
\end{itemize}
\par}
\end{keybox}

\tableofcontents
\newpage

\section{The summit claim}\label{sec:summit-claim}

\subsection{The central statement}

The program distilled in this paper is centered on the following claim.

\begin{theorem}[Reflective Non-Exhaustion Theorem]\label{thm:rne}
Let a \emph{reflective certification architecture} consist of:
\begin{enumerate}[label=(\roman*)]
  \item a bare certification carrier $B$;
  \item an enriched realization carrier $E$;
  \item a sound comparison map $\pi : E \to B$;
  \item a distinguished canonical region $C \subseteq B$ singled out by the theory's
    certification rules.
\end{enumerate}
Then the completed formal instances studied in this program support the following
uniform law---a dichotomy between total collapse along $\pi$ and genuine
non-exhaustion:
\begin{enumerate}[label=(\alph*)]
  \item \emph{(collapse regime)} The enriched layer carries no nontrivial residue
    above certification for the comparison problem governed by $\pi$.
  \item \emph{(non-exhaustion regime)} Canonical certification collapses inside $C$,
    the realization layer remains nontrivial above that collapse, and the
    nontriviality is organized by non-injectivity of $\pi$, by nontrivial fibers
    $\pi^{-1}(b)$ for $b \in C$, and by explicit obstruction or section laws governing
    when collapse can or cannot be lifted.
\end{enumerate}
In the non-exhaustion regime, canonical certification does not exhaust realization.
The suite's completed formal discharges instantiate this non-exhaustion regime in
distinct mathematical domains.
\end{theorem}

\noindent
The completed discharges justify reading this law as the governing synthesis of the
current program; a strongest universal form over the broadest abstract class remains
future work (\cref{sec:levels}).

\noindent
Here \emph{reflective} means that the architecture contains both an official
certification layer and a structured realization layer connected by an explicit
comparison map, so that the theory can internally track the relation between what
is certified and how that certification is realized.

\noindent
In the non-exhaustion regime, the residue does not live on the bare carrier itself
but in the fibers of $\pi$; liftability is governed by obstruction theory, not by
certificate theory alone. Across the suite, this is one shared architecture, not
family resemblance.

\noindent
Theorem~\ref{thm:rne} is a summit-level synthesis statement. Its formal instances are
proved in the companion papers and Lean modules cited below; this paper does not ask
the reader to accept it as a slogan detached from formal mathematics.

\subsection{Public paraphrase}

A concise public rendering is:

\begin{quote}
\emph{A system's canonical certificate need not fully contain the structure by
which that certificate is realized.}
\end{quote}

A stronger paraphrase:

\begin{quote}
\emph{Canonical certification does not, in general, exhaust realized structure.}
\end{quote}

\subsection{What this paper is for}

This paper is not the place where the detailed Lean proofs live; nor is it a
popular essay. Its role is to do four things that none of the companion papers,
taken alone, can do as sharply:

\begin{enumerate}[label=(\arabic*)]
  \item state the permanent theorem-level idea in one place;
  \item show exactly which completed results already support it;
  \item separate it from nearby but different ideas (incompleteness, generic proof
    relevance, informal rhetoric);
  \item explain why the architecture is a reusable research instrument rather than
    a one-off observation.
\end{enumerate}

\subsection{Program route labels}\label{sec:route-labels}

The suite groups its formal work under named \emph{routes} (navigation labels, not
routing tables or inference rules). \textbf{Route~C} is the headline reflective
non-exhaustion claim; \textbf{Routes~A} and \textbf{B} cover infrastructure lemmas
and the algebra/topology/arithmetic general-method papers; \textbf{Route~D} is the
logic/computability discharge---halting and Rice scaffolding in Mathlib's
\texttt{ComputablePred} layer (Lean \texttt{RouteD/}). This vocabulary is fixed here
so that later mentions of Route~D are not unexplained forward references.

\section{Three levels of claim}\label{sec:levels}

This paper distinguishes three levels; readers should not conflate them.

\paragraph{Instance level.}
The companion papers prove specific theorems on named carriers in Lean.

\paragraph{Synthesis level.}
This paper identifies the common architecture those theorems instantiate and states
it as the summit thesis (Theorem~\ref{thm:rne}).

\paragraph{Frontier level.}
A stronger universal inevitability theorem over the broadest abstract class remains
future work.

\section{Why this counts as theorem-level synthesis}\label{sec:theorem-synthesis}

The summit claim is not a narrative imposed after the fact. It is theorem-level
because:
\begin{itemize}
  \item \textbf{Named formal instances.} Companion papers and Lean modules prove
    concrete theorems on named carriers (zero \texttt{sorry} in the relevant
    modules).
  \item \textbf{Shared comparison structure.} The same pattern---bare layer, enriched
    layer, comparison map $\pi : E \to B$---is instantiated across internal, benchmark,
    algebraic, arithmetic, topological, and computability-adjacent settings.
  \item \textbf{Fiber and obstruction law.} Residual structure is located in fibers
    and controlled by obstruction in each discharge, not by certificate rhetoric
    alone.
\end{itemize}
Theorem~\ref{thm:rne} is the explicit synthesis statement these instances support.

\section{What is actually proved in the suite}\label{sec:proved-suite}

The summit claim is supported by several completed theorem families, all formalized
in Lean~4 with zero \texttt{sorry} in the relevant modules.

\subsection{Internal origin theorem}

The flagship Infinity Compression paper~\cite{SpivackIC} proves the core
certification/realization split on named carriers internal to the library:

\begin{enumerate}[label=(\arabic*)]
  \item canonical collapse at the bare certification carrier;
  \item a reflective split theorem on enriched witnesses;
  \item strict refinement via a typed forgetful map;
  \item nontriviality of the canonical fiber;
  \item injective role coordinates on a standard fiber patch.
\end{enumerate}

This is the \emph{internal origin theorem}: the place where the architecture first
appears not as vocabulary but as a proved, load-bearing structure.

\subsection{External transfer}

The validation paper~\cite{SpivackValidation} proves that the same packaging
exports to twelve independent mathematical benchmark families: quotients,
products, dependent sums, coproduct discriminants, subtypes, orbit quotients,
ideal quotients, pushouts, classifiers, localization, pullbacks, and a
non-surjective control case.

The role of that paper is not theorem-level domain leverage. Its role is
transferability: it demonstrates that the architecture is not an idiosyncratic
artifact of one flagship library.

\subsection{Algebraic discharge}

The group extension paper~\cite{SpivackFiber} proves the central algebraic
obstruction law:
\[
\text{splits} \iff \text{trivial cocycle}.
\]
It also proves the section cocycle identity, local-to-global obstruction
principles, and associated fiber structure.

This is a genuine theorem-level discharge in classical mathematics: the
architecture is not merely organizational there, but mathematically load-bearing.

\subsection{Arithmetic discharge}

The same paper and associated Lean modules show that embedding problems are
equivalent to extension splitting problems and hence to cocycle triviality:
\[
\text{solvable embedding problem}
\iff
\text{splitting}
\iff
\text{trivial cocycle}.
\]
This supplies the arithmetic face of the program, connecting the architecture to
inverse Galois-style lifting problems.

\subsection{Topological discharge}

The Quillen paper~\cite{SpivackQuillen} proves Quillen's Theorem~A for Galois
connections at the simplicial/nerve level, using contractible fibers and
homotopy-equivalence data. This is the topological instance of the same
architecture: fibers, explicit comparison maps, and a theorem-level consequence.

\subsection{Computability anchor}

The summit-supporting Lean modules also contain a logic/computability anchor:
halting is not capturable by the relevant certification layer, and Rice-style
phenomena block certification of nontrivial semantic properties in the
\texttt{ComputablePred} setting. This does not mean the program is ``just
G\"odel again,'' but it ensures the summit architecture touches standard
non-certifiability barriers in classical computability theory.

\subsection{Summary of support}

Taken together, the suite already contains:
\begin{center}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Support type} & \textbf{Completed instance} \\
\midrule
Internal origin & certification vs.\ realization split in IC \\
Transferability & twelve external benchmark families \\
Algebraic discharge & group extension splitting criterion \\
Arithmetic discharge & embedding problems and tower bridge \\
Topological discharge & Quillen A for Galois connections \\
Computability anchor & halting / Rice non-certifiability layer \\
\bottomrule
\end{tabular}
\end{center}

That support is sufficient to justify a summit paper: the summit claim is no
longer a program aspiration but a synthesis of completed formal discharges.

\section{The architecture itself}\label{sec:architecture}

\subsection{Certification}

A \emph{certification layer} is a carrier of theorem-level or officially packaged
invariants---the layer at which a theory certifies an object, witness, or verdict.
Its defining feature here is that it may \emph{collapse}: many legitimate routes may
project to the same canonical certification object.

\subsection{Realization}

A \emph{realization layer} is an enriched carrier above the certificate. It tracks
the structure by which the certificate is obtained or instantiated: sections,
role assignments, dependency shapes, enriched reflective witnesses, splitting
data, chosen representatives, homotopy witnesses, and similar objects.

Realization is not decoration. It becomes mathematically serious exactly when the
comparison map from realization to certification is not injective.

\subsection{Comparison map}

The \emph{comparison map} or \emph{forgetful map}
\[
\pi : E \to B
\]
forgets enriched structure and records only the bare certified image.

This map is the central organizing object. Once $\pi$ is explicit, one can ask:
\begin{enumerate}[label=(\alph*)]
  \item Is $\pi$ surjective?
  \item Is $\pi$ injective?
  \item What do the fibers $\pi^{-1}(b)$ look like?
  \item Are there canonical sections or right inverses?
  \item What obstruction governs lifting from the bare layer back upstairs?
\end{enumerate}

\subsection{Fibers}

The fiber over $b \in B$ is
\[
\Fib(b) := \pi^{-1}(b).
\]
When certification collapses, fibers are where residual structure lives---the usual
fiber picture in topology~\cite{Steenrod1951}. Multiple points in $\Fib(b)$ mean bare
certification does not fix realization uniquely.

\subsection{Obstruction}

An \emph{obstruction law} is the theorem governing when one may lift from bare to
enriched structure. Lifting and obstruction in the sense of fiber bundles and
homotopy theory are standard~\cite{Hatcher2002,Steenrod1951}. In different domains
this obstruction appears in different mathematical clothing:
\begin{itemize}
  \item a cocycle in group extensions,
  \item fiber contractibility in Quillen-style comparison,
  \item existence or failure of sections and right inverses in benchmark families,
  \item non-certifiability in computability.
\end{itemize}

The summit claim is that these phenomena instantiate one
collapse/refinement/fiber/obstruction architecture, not unrelated accidents.

\section{The internal origin theorem}\label{sec:internal-origin}

The flagship result~\cite{SpivackIC} is the place where the summit architecture is
first proved, rather than merely observed.

\subsection{What happens internally}

At the bare carrier, the standard Phase-2 extraction collapses to a unique canonical
certificate. At the enriched carrier, distinct reflective witnesses remain distinct.
A typed forgetful map from enriched reflective splits to canonical bare certificates
is sound but not injective. The canonical fiber above the bare basepoint is
nontrivial, and role data injects along a standard subfamily.

\subsection{Why that matters}

This gives the program its sharpest internal lesson:

\begin{quote}
Once collapse at the bare carrier is proved, comparison questions that still matter
must move upstairs to enriched carriers and fibers.
\end{quote}

That is a mathematical, not stylistic, conclusion. The internal origin theorem is
what turns the rest of the suite into a coherent program rather than a collection
of analogies.

\section{Transferability}\label{sec:transferability}

The benchmark survey~\cite{SpivackValidation} proves that the architecture exports
to many ordinary mathematical constructions.

\subsection{Why transfer matters}

Without transfer, the flagship theorem could be dismissed as an artifact of an
idiosyncratic internal formalism. The tranche survey blocks that dismissal. It
shows that the same language of:
\[
\text{bare layer} \quad \text{vs.} \quad \text{enriched layer} \quad \text{vs.} \quad
\text{fiber / section / right inverse}
\]
is not local to one formal development.

\subsection{What transfer does not claim}

Transferability is not itself the summit theorem. Nor does it prove new deep
classical results in each benchmark family. Its function is narrower and
essential: it validates the architecture as a reusable interface.

\section{Algebraic and arithmetic discharge}\label{sec:alg-arith}

\subsection{Group extensions}

The group extension theorem
\[
\text{splits} \iff \text{trivial cocycle}
\]
is a model example of the summit architecture:
\begin{itemize}
  \item the bare layer is set-theoretic section existence, which always holds;
  \item the enriched layer is homomorphic splitting;
  \item the comparison is the passage from structured splitting data to bare section data;
  \item the obstruction is the section 2-cocycle.
\end{itemize}

This is exactly the sort of result that makes the architecture mathematically
load-bearing rather than merely descriptive.

\subsection{Embedding problems}

Embedding problems sharpen the picture:
\[
\text{solvable} \iff \text{splits} \iff \text{trivial cocycle}.
\]
Here the same architecture governs an arithmetic lifting problem. Bare data does
not decide liftability by itself; the residue sits in the obstruction.

\subsection{Arithmetic significance}

This is important for the summit claim because it shows that the architecture has
teeth in a domain that is not just internal formal theory and not just generic
category packaging. It reaches into a recognized arithmetic problem class.

\section{Topological discharge}\label{sec:topology}

The Quillen paper~\cite{SpivackQuillen} provides the topological discharge.

\subsection{What happens there}

For Galois connections, the relevant fibers are principal down-sets with terminal
or maximal structure, hence contractible nerves. This gives the comparison theorem:
the induced nerve maps carry explicit homotopy-equivalence data.

\subsection{Architectural reading}

In that setting:
\begin{itemize}
  \item the comparison map is the induced nerve map;
  \item the fibers encode the local comparison geometry;
  \item contractibility is the obstruction-theoretic condition that trivializes
    the comparison problem.
\end{itemize}

This is a very different mathematical domain from group extensions, yet the same
architecture reappears.

\section{Computability anchor}\label{sec:computability}

\textbf{Route~D} (\cref{sec:route-labels}) is the computability discharge: halting and
Rice-style facts in Mathlib's \texttt{ComputablePred} layer, formalized in
\texttt{RouteD/}. They show semantic phenomena that need not be fully capturable by
the relevant certification apparatus, with classical landmarks in incompleteness and
undecidability~\cite{Godel1931,Turing1936,Rice1953}. This prevents the summit
architecture from being mistaken for a vague philosophical analogue of G\"odel's
theorem~\cite{Godel1931}.

In parallel with the other domains, the bare layer is the certifiable predicate
interface; the enriched layer is semantic behavior of computations; the comparison is
certification through \texttt{ComputablePred}; the obstruction is classical
undecidability (halting / Rice).

\subsection{What is not claimed}

The summit theorem is not G\"odel's incompleteness theorem~\cite{Godel1931}, nor is
the present program claiming to subsume all of classical recursion theory. The
anchor is narrower: the suite's language touches standard impossibility terrain in
an honest, formal way.

\section{Why this is not just G\"odel}\label{sec:not-godel}

This distinction is central.

\subsection{G\"odel's contribution}

G\"odel shows that sufficiently rich formal systems face unavoidable limits:
undecidable statements, incompleteness, and non-recursive truth phenomena~\cite{Godel1931}.

\subsection{This program's contribution}

The present program shows something structurally different:
\begin{enumerate}[label=(\arabic*)]
  \item where collapse occurs;
  \item where it provably does not occur;
  \item how the non-collapse is organized.
\end{enumerate}

That is a positive architecture theorem, not merely a barrier theorem.

\subsection{The sharp difference}

G\"odel's theorem says, roughly, that certain truths outrun formal provability.
The summit architecture addresses a different question: where certification
collapses at one carrier, realized structure may still be nontrivial upstairs, with
that nontriviality organized by fibers and obstructions (Theorem~\ref{thm:rne},
non-exhaustion regime).

The two ideas are related in spirit once \textbf{Route~D} (halting and Rice in
\texttt{ComputablePred}) is in play, but they are not the same theorem.

\section{Why this is not merely proof relevance in type theory}\label{sec:not-dtt}

A common dismissal would be: of course enriched objects exist in dependent type
theory; proof relevance is built in.

That is not the result.

\subsection{What would be trivial}

It is trivial that richer carriers can be defined in DTT. It is trivial that one
can invent a forgetful map from a rich type to a poor one.

\subsection{What is nontrivial here}

What is nontrivial is that the library proves:
\begin{enumerate}[label=(\arabic*)]
  \item collapse at one named carrier;
  \item non-collapse at another named carrier;
  \item non-injectivity of the explicit comparison map;
  \item nontriviality and partial classification of the corresponding fibers;
  \item parallel obstruction laws in external mathematics.
\end{enumerate}

The summit claim is therefore not ``proof relevance exists.'' It is:
\begin{quote}
\emph{this is the correct theorem-level architecture for a broad and growing class of
comparison problems.}
\end{quote}

\section{Why this changes mathematical practice}\label{sec:practice}

The strongest reason to take the summit claim seriously is methodological.

\subsection{The reusable recipe}

The architecture gives a reusable research workflow:

\begin{enumerate}[label=(\arabic*)]
  \item identify the bare certificate carrier $B$;
  \item identify an enriched carrier $E$;
  \item define the comparison map $\pi : E \to B$;
  \item prove whether collapse occurs at $B$;
  \item test injectivity or non-injectivity of $\pi$;
  \item if non-injective, study the fibers $\Fib(b)$;
  \item identify sections, right inverses, or obstruction laws governing liftability.
\end{enumerate}

\subsection{Why this matters}

This turns vague ``more structure upstairs'' claims into a precise program: it tells
researchers what to formalize and where. The architecture is therefore a
\emph{research instrument}---a stable way to formulate and attack comparison
problems where theorem-level invariants do not exhaust realizing structure.

\section{Cross-domain dictionary}\label{sec:dictionary}

The following table is a \emph{heuristic} index: it is not a formal theorem, only a
compact navigation aid.

\begin{center}
\resizebox{\textwidth}{!}{%
\footnotesize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{0.16\textwidth}>{\raggedright\arraybackslash}p{0.20\textwidth}>{\raggedright\arraybackslash}p{0.20\textwidth}>{\raggedright\arraybackslash}p{0.19\textwidth}>{\raggedright\arraybackslash}p{0.21\textwidth}@{}}
\toprule
\textbf{Domain} & \textbf{Bare layer} & \textbf{Enriched layer} & \textbf{Comparison map} & \textbf{Fiber / obstruction} \\
\midrule
IC flagship & canonical bare certification & enriched reflective split & typed forgetful map $\pi_A$ & nontrivial canonical fiber \\
Benchmarks & baseline representability & structured witness bundle & projection / forgetting & fibers, sections, right inverses \\
Group extensions & set-theoretic section layer & splitting data & forgetting homomorphic structure & cocycle obstruction \\
Embedding problems & bare target quotient data & lifting / solution data & solution-to-extension comparison & solvable iff trivial cocycle \\
Quillen GC & order-complex / nerve image & homotopy witness data & induced nerve maps & contractible fibers \\
Computability & certifiable predicate layer & realized semantic behavior & certification comparison & halting / Rice barriers \\
\bottomrule
\end{tabular}%
}
\end{center}

It is nonetheless the best concise representation of why the summit architecture is
plausible as a unifying language.

\section{Exact scope}\label{sec:scope}

The summit claim is strong, but it is not unlimited. The following paragraphs make
the \emph{synthesis}-level claim of \cref{sec:levels} exact and separate it from the
\emph{frontier} level.

\subsection{What is claimed}

We claim that:
\begin{enumerate}[label=(\arabic*)]
  \item the collapse/refinement/fiber/obstruction architecture is formally realized
    in the completed companion work;
  \item it transfers across multiple external families;
  \item it yields theorem-level leverage in algebra, arithmetic, topology, and
    computability-adjacent settings;
  \item it is therefore a plausible permanent language for a broad class of
    comparison problems.
\end{enumerate}

\subsection{What is not claimed}

We do \emph{not} claim:
\begin{enumerate}[label=(\arabic*)]
  \item a single current Lean theorem quantifying over all domains at once;
  \item that every formal system must instantiate this exact architecture;
  \item that the current program has already achieved universal adoption or full
    G\"odel-level social impact.
\end{enumerate}

\paragraph{Outside scope.}
Possible speculative applications to self-modeling or phenomenological systems lie outside
the formal scope of this paper; they are not part of the proved theorem content.

\subsection{Why this is enough}

These non-claims do not weaken the summit paper's purpose. A summit statement
should mark the strongest theorem-level language already earned by the completed
program, not pretend to have finished all mathematically natural extensions.

\section{Open frontier}\label{sec:frontier}

The next frontier is clear.

\subsection{Mathematical extensions}

The strongest natural next steps are:
\begin{enumerate}[label=(\arabic*)]
  \item fuller universal quantification over a broad abstract architecture class;
  \item stronger fiber-classification theorems;
  \item completion of the $H^1/H^2$ bridge in Mathlib;
  \item deeper arithmetic realization through concrete Galois tower infrastructure;
  \item general Quillen Theorem~A beyond the Galois-connection case.
\end{enumerate}

\subsection{Program-level extensions}

At the level of program closure, the remaining work is not to invent an entirely
new idea, but to strengthen and universalize what is already visible:
\begin{itemize}
  \item sharper inevitability theorems,
  \item stronger unified statements,
  \item broader accepted examples,
  \item publication and adoption.
\end{itemize}

\section{Conclusion}\label{sec:conclusion}

The program's permanent theorem-level claim is now stated in one place:
Theorem~\ref{thm:rne}.

It is not a slogan about proof objects, a rebranding of G\"odelian incompleteness,
or an artifact of one library. Six completed discharges support it (internal origin,
transferability, algebra, arithmetic, topology, computability).

\begin{quote}
\emph{Collapse downstairs, structured comparison upstairs, and obstruction in the
fibers is a real, reusable theorem-level architecture.}
\end{quote}

What survives collapse has a precise mathematical home in the fibers of comparison,
governed by obstruction. The right language is certification together with the geometry of what
certification forgets.

\bibliographystyle{alpha}
\begin{thebibliography}{99}

\bibitem{SpivackIC}
N.~Spivack.
\newblock \emph{Canonical Certification Does Not Exhaust Reflective Structure:
  Reflective Split, Strict Refinement, and Fiber Structure in Infinity Compression}.
\newblock Flagship manuscript, 2026.

\bibitem{SpivackValidation}
N.~Spivack.
\newblock \emph{External Validation of a Positive-Closure Proof Architecture:
  A Multi-Tranche Survey}.
\newblock Validation manuscript, 2026.

\bibitem{SpivackFiber}
N.~Spivack.
\newblock \emph{Fiber Architecture for Group Extensions in Lean~4: Cocycles, Splitting,
  and the Cohomological Bridge}.
\newblock General-method manuscript, 2026.

\bibitem{SpivackQuillen}
N.~Spivack.
\newblock \emph{Quillen's Theorem~A for Galois Connections: A Machine-Checked
  Formalization in Lean~4}.
\newblock General-method manuscript, 2026.

\bibitem{Godel1931}
K.~G\"odel.
\newblock {\"U}ber formal unentscheidbare S{\"a}tze der \emph{Principia Mathematica} und
  verwandter Systeme~I.
\newblock \emph{Monatshefte f\"ur Mathematik und Physik} \textbf{38} (1931), 173--198.

\bibitem{Turing1936}
A.~M.~Turing.
\newblock On computable numbers, with an application to the Entscheidungsproblem.
\newblock \emph{Proc.\ London Math.\ Soc.} (2) \textbf{42} (1936), 230--265.

\bibitem{Rice1953}
H.~G.~Rice.
\newblock Classes of recursively enumerable sets and their decision problems.
\newblock \emph{Trans.\ Amer.\ Math.\ Soc.} \textbf{74} (1953), 358--366.

\bibitem{Steenrod1951}
N.~Steenrod.
\newblock \emph{The Topology of Fibre Bundles}.
\newblock Princeton University Press, 1951.

\bibitem{Hatcher2002}
A.~Hatcher.
\newblock \emph{Algebraic Topology}.
\newblock Cambridge University Press, 2002.

\end{thebibliography}

\end{document}