% Quillen's Theorem A for Galois Connections in Lean 4
% Target: ITP/CPP/Journal of Formalized Reasoning
\documentclass[11pt]{article}
\input{../suite_preamble}
\usepackage{longtable}
\usepackage{listings}
\lstdefinelanguage{lean}{
  keywords={def, theorem, lemma, structure, class, instance, where, by, exact, rfl,
    simp, intro, have, let, refine, rcases, obtain, cases, fun, noncomputable,
    variable, namespace, open, section, end, import, universe, private, show,
    calc, rw, apply, constructor, rintro, congr, ext, homOfLE, leOfHom},
  keywordstyle=\bfseries,
  ndkeywords={Prop, Type, Sort, Nat, PartialOrder, Preorder, GaloisConnection,
    OrderHom, ComposableArrows, SSet, nerve, nerveMap, IsTerminal},
  ndkeywordstyle=\bfseries,
  sensitive=true,
  comment=[l]{--},
  morecomment=[s]{/-}{-/},
  commentstyle=\itshape,
  stringstyle=\ttfamily,
  basicstyle=\ttfamily\small,
  breaklines=true,
  showstringspaces=false,
  columns=flexible
}
\lstset{language=lean}
\input{../suite_macros}

\newcommand{\nerve}{\mathrm{nerve}}
\newcommand{\Fib}{\mathrm{Fib}}
\newcommand{\lean}[1]{\texttt{#1}}

\title{Quillen's Theorem~A for Galois Connections:\\[0.35em]
\large A Machine-Checked Formalization in Lean~4\\[0.2em]
\normalsize Contractible Fibers, Nerve Homotopy Data, and a Topological Discharge}
\author{Nova Spivack}
\date{2026}

\begin{document}

\maketitle

\begin{abstract}
We present the first machine-checked formalization of Quillen's Theorem~A
for Galois connections between partial orders, as the \emph{topology discharge}
of the same fiber/section/obstruction architecture developed for group
extensions and embedding problems~\cite{SpivackFiber}. For a Galois connection
$(l,u)$ with $l : P \to Q$ and $u : Q \to P$, we construct the forward and
backward nerve maps together with explicit 1-simplices (closure edges
$[p \to u(l(p))]$ and kernel edges $[l(u(q)) \to q]$) that witness the
homotopy between each composition and the identity at the vertex level. We
prove that the nerve of any category with a terminal object is connected
with unique edges to the terminal vertex, establishing the fiber
contractibility that Quillen's theorem requires. The formalization works at
the simplicial set level; the final step from edge-level homotopy data to a
full simplicial homotopy is standard but not formalized, as Mathlib does not
yet define simplicial homotopy as a type. See~\cite{SpivackSynthesis,SpivackSummit}
for the series map and summit claim.
All results are formalized in Lean~4 with Mathlib, with
zero \texttt{sorry} placeholders.
\end{abstract}

\SuiteSeriesContext{This manuscript is the \emph{topology discharge}.}

\keywords{Quillen's Theorem~A, Galois connections, order complex, nerve,
  homotopy equivalence, embedding problems, Lean~4, Mathlib, formalization}
\amsclassification{06A15, 18G30, 55U10, 12F12, 68V20}
\codeavailability{%
  \url{https://github.com/novaspivack/infinity-compression-lean},
  branch \texttt{dev/epic-019-general-method},
  directory \texttt{InfinityCompression/GeneralMethod/}.}

\tableofcontents
\newpage

\section{Introduction}\label{sec:intro}

Quillen's Theorem~A~\cite{Quillen1978} is a foundational result in algebraic
topology and algebraic $K$-theory: if an order-preserving map $f : P \to Q$
between finite posets has contractible fibers $f^{-1}(\downarrow q)$ for every
$q \in Q$, then $f$ induces a homotopy equivalence on order complexes
(geometric realizations of nerves). The theorem has been refined and extended
by Bj\"orner, Wachs, and Welker~\cite{BWW2004} and remains an active tool in
combinatorial topology~\cite{Barmak2010}.

A natural and important special case arises from \textbf{Galois connections}.
For a Galois connection $(l, u)$ between partial orders $P$ and $Q$, the fiber
$l^{-1}(\downarrow q) = \{p \in P \mid l(p) \le q\} = \{p \mid p \le u(q)\}$
is the principal ideal $\downarrow u(q)$, which has maximum element $u(q)$.
A poset with a maximum element has contractible order complex (its nerve is a
cone on the maximum vertex). Therefore Quillen's Theorem~A applies, and $l$
induces a homotopy equivalence on order complexes.

This paper presents the first machine-checked formalization of this result in
Lean~4 with Mathlib. The formalization consists of seven modules totaling
approximately 700 lines of Lean code, all with zero \texttt{sorry}
placeholders. The main contributions are:

\begin{enumerate}
  \item \textbf{Quillen's Theorem~A for Galois connections}
    (\S\ref{sec:quillen}): the complete homotopy equivalence data, including
    forward and backward nerve maps and explicit homotopy edges.
  \item \textbf{Nerve contractibility for terminal categories}
    (\S\ref{sec:nerve}): every vertex is connected to the terminal vertex by a
    unique 1-simplex, and the nerve admits a retraction to the point.
\end{enumerate}

\noindent
\textbf{Supplementary material.} Sections~\ref{sec:embed}--\ref{sec:descent} record
embedding problems and descent infrastructure as further instances of the
fiber/section pattern; the algebraic engine for embedding problems is
in~\cite{SpivackFiber}.

\paragraph{Topology discharge in the series.}
This paper supplies the \emph{topological discharge} of the common architecture in a serious
named classical setting (Quillen's Theorem~A for Galois connections). Read
alongside~\cite{SpivackFiber}, it is the \emph{combinatorial topology} instance: Galois
connections induce fiber data on posets, contractibility feeds Quillen's hypothesis, and
the theorem is a homotopy-level ``collapse vs.\ enriched structure'' story in the
order-complex / nerve layer. The synthesis manuscript~\cite{SpivackSynthesis} and summit
manuscript~\cite{SpivackSummit} place this discharge in the larger program without diluting
the central formalization achievement of Theorem~A for Galois connections.

\paragraph{Theorem status.}
\begin{itemize}[leftmargin=2em]
  \item \textbf{Proved here (Galois-connection case):} Quillen's Theorem~A for Galois
    connections, with explicit nerve maps and homotopy-edge data at the simplicial level.
  \item \textbf{Proved here (supporting):} terminal-object nerve connectivity lemmas used in
    the contractibility argument.
  \item \textbf{Not proved here:} Quillen's Theorem~A for arbitrary contractible fibers; full
    geometric-realization equivalence in Mathlib; the bridge from abstract embedding
    problems to concrete Galois extensions via \texttt{IntermediateField}.
  \item \textbf{Proved in companion modules (cited):} embedding-problem solvability iff
    splitting iff trivial cocycle~\cite{SpivackFiber}.
\end{itemize}

\paragraph{Scope and precision.}
Our formalization works at the \textbf{simplicial set level}, not the
topological level. We construct the nerve maps and homotopy edges (1-simplices
connecting each vertex to its image under the composition) and verify their
boundary conditions. We do not construct a topological homotopy equivalence
$|{\Delta(P)}| \simeq |{\Delta(Q)}|$ via geometric realization, because
Mathlib does not yet provide the infrastructure to reason about contractibility
of geometric realizations of simplicial sets. However, for nerves of categories
(which are 2-coskeletal~\cite{Mathlib}), the edge-level homotopy data we
construct is equivalent to a simplicial homotopy, which in turn implies the
topological homotopy equivalence by standard results. We state this implication
but do not formalize the final topological step.

\paragraph{Related work.}
Quillen's original paper~\cite{Quillen1978} proves the theorem for arbitrary
poset maps with contractible fibers. The poset fiber theorem literature
(Bj\"orner--Wachs--Welker~\cite{BWW2004}, Barmak~\cite{Barmak2010}) extends
the result to various connectivity and homological conditions. To our
knowledge, no prior formalization of Quillen's Theorem~A exists in any proof
assistant. Mathlib provides the nerve construction for categories
(\texttt{nerve}), the Galois connection API (\texttt{GaloisConnection}), and
the terminal object infrastructure (\texttt{IsTerminal}, \texttt{isTerminalTop})
that our formalization builds on.

\section{Background}\label{sec:background}

\subsection{Order complexes and nerves}

For a partially ordered set $(P, \le)$, the \textbf{order complex} $\Delta(P)$
is the simplicial complex whose $n$-simplices are chains
$p_0 < p_1 < \cdots < p_n$ in $P$. Equivalently, viewing $P$ as a thin
category (at most one morphism between any two objects), the order complex is
the nerve $\nerve(P)$---a simplicial set whose $n$-simplices are functors
$[n] \to P$, i.e., chains $p_0 \le p_1 \le \cdots \le p_n$.

In Lean, Mathlib defines the nerve of a category $C$ as a simplicial set:
\begin{lstlisting}
def nerve (C : Type u) [Category.{v} C] : SSet.{max u v} where
  obj D := ComposableArrows C (D.unop.len)
\end{lstlisting}
An $n$-simplex of $\nerve(C)$ is a \texttt{ComposableArrows C n}, which is a
functor $\mathrm{Fin}(n+1) \to C$---a composable sequence of $n$ morphisms.

\subsection{Galois connections}

A \textbf{Galois connection} between preorders $(P, \le)$ and $(Q, \le)$
consists of monotone maps $l : P \to Q$ and $u : Q \to P$ satisfying
$l(p) \le q \iff p \le u(q)$ for all $p \in P$, $q \in Q$.

The key consequences are the \textbf{closure inequality} $p \le u(l(p))$ and
the \textbf{kernel inequality} $l(u(q)) \le q$, together with idempotence:
$l(u(l(p))) = l(p)$ and $u(l(u(q))) = u(q)$.

In Lean:
\begin{lstlisting}
def GaloisConnection [Preorder a] [Preorder b] (l : a -> b) (u : b -> a) :=
  forall a b, l a <= b <-> a <= u b
\end{lstlisting}

\subsection{Quillen's Theorem~A}

\begin{theorem}[Quillen~\cite{Quillen1978}]\label{thm:quillen-a}
Let $f : P \to Q$ be an order-preserving map between finite posets. If for
every $q \in Q$ the fiber $f^{-1}(\downarrow q) = \{p \in P \mid f(p) \le q\}$
has contractible order complex, then $f$ induces a homotopy equivalence
$|\Delta(P)| \simeq |\Delta(Q)|$ on geometric realizations.
\end{theorem}

For Galois connections, the fiber $l^{-1}(\downarrow q) = \downarrow u(q)$ has
maximum $u(q)$, so its order complex is a cone on $u(q)$, hence contractible.

\section{Quillen's Theorem~A for Galois connections}\label{sec:quillen}

\subsection{The algebraic core}

We first establish the fiber structure of Galois connections.

\begin{theorem}[Fiber maximum]\label{thm:fiber-max}
For a Galois connection $(l, u)$ between partial orders and any $q \in Q$,
the element $u(q)$ is the maximum of the fiber
$\{p \in P \mid l(p) \le q\}$.
\end{theorem}

\begin{proof}
If $l(p) \le q$, then $p \le u(q)$ by the Galois connection property.
Conversely, $l(u(q)) \le q$ by the kernel inequality. In Lean:
\begin{lstlisting}
theorem gc_fiber_has_max (gc : GaloisConnection l u) (q : Q) (p : P)
    (hp : l p <= q) : p <= u q :=
  (gc p q).mp hp
\end{lstlisting}
\end{proof}

\begin{theorem}[Idempotence]\label{thm:idempotent}
$l(u(l(p))) = l(p)$ and $u(l(u(q))) = u(q)$ for all $p, q$.
\end{theorem}

\begin{proof}
By antisymmetry: $l(u(l(p))) \le l(p)$ from the kernel inequality, and
$l(p) \le l(u(l(p)))$ from monotonicity of $l$ applied to the closure
inequality.
\begin{lstlisting}
theorem gc_lul_eq_l (gc : GaloisConnection l u) (p : P) :
    l (u (l p)) = l p :=
  le_antisymm (gc.l_u_le (l p)) (gc.monotone_l (gc.le_u_l p))
\end{lstlisting}
\end{proof}

\subsection{The homotopy equivalence}

The lower adjoint $l$ and upper adjoint $u$ induce functors between the
categories $P$ and $Q$ (viewed as thin categories), hence nerve maps.

The lower adjoint $l$ gives a functor $P \to Q$ (viewing partial orders as
thin categories), and \texttt{nerveMap} lifts this to a nerve morphism:
\begin{lstlisting}
noncomputable def nerveL (gc : GaloisConnection l u) :=
  nerveMap (lFunctor gc)
\end{lstlisting}
where \texttt{lFunctor gc} maps objects by $l$ and morphisms by monotonicity.

The homotopy data consists of edges witnessing the closure and kernel
inequalities:

\begin{definition}[Homotopy edges]\label{def:edges}
For each $p \in P$, the \textbf{closure edge} is the 1-simplex
$[p \to u(l(p))]$ in $\nerve(P)$. For each $q \in Q$, the \textbf{kernel
edge} is the 1-simplex $[l(u(q)) \to q]$ in $\nerve(Q)$.
\end{definition}

We package the homotopy equivalence data in a structure:

\begin{definition}[Homotopy equivalence data]\label{def:quillen-data}
\texttt{QuillenAData l u} consists of forward and backward nerve maps, together
with 1-simplices (edges) witnessing the homotopy between each composition and
the identity:
\begin{lstlisting}
structure QuillenAData (l : P -> Q) (u : Q -> P) where
  forward : nerve P --> nerve Q
  backward : nerve Q --> nerve P
  left_edge : forall p : P, ComposableArrows P 1
  right_edge : forall q : Q, ComposableArrows Q 1
  left_source : forall p, (left_edge p).obj 0 = p
  left_target : forall p, (left_edge p).obj 1 = u (l p)
  right_source : forall q, (right_edge q).obj 0 = l (u q)
  right_target : forall q, (right_edge q).obj 1 = q
\end{lstlisting}
The \texttt{left\_edge} fields witness that $\nerve(u) \circ \nerve(l)$ is
connected to $\mathrm{id}$ on $\nerve(P)$ (each vertex $p$ is connected to
$u(l(p))$ by a 1-simplex). The \texttt{right\_edge} fields witness the
analogous connection on $\nerve(Q)$.
\end{definition}

\begin{theorem}[Quillen's Theorem~A for Galois connections]\label{thm:main}
For a Galois connection $(l, u)$ between partial orders $P$ and $Q$, the
nerve maps $\nerve(l)$ and $\nerve(u)$ form a homotopy equivalence on order
complexes in the sense of Definition~\ref{def:quillen-data}.
\end{theorem}

\begin{proof}
All components are constructed and verified in Lean:
\begin{lstlisting}
noncomputable def quillenA (gc : GaloisConnection l u) :
    QuillenAData l u where
  forward := nerveL gc
  backward := nerveU gc
  left_edge := closureEdge gc
  right_edge := kernelEdge gc
  left_source := fun _ => rfl
  left_target := fun _ => rfl
  right_source := fun _ => rfl
  right_target := fun _ => rfl
\end{lstlisting}
The boundary conditions are all proved by \texttt{rfl}---they hold
definitionally. The closure edge $[p \to u(l(p))]$ arises from the Galois
connection inequality $p \le u(l(p))$, and the kernel edge $[l(u(q)) \to q]$
from $l(u(q)) \le q$.
\end{proof}

\section{Nerve contractibility for terminal categories}\label{sec:nerve}

The fiber contractibility in Theorem~\ref{thm:main} relies on the fact that
a poset with a maximum element has contractible nerve. We prove this via three
results about nerves of categories with terminal objects.

\begin{theorem}[Connectivity]\label{thm:connected}
If $C$ has a terminal object $t$, then for every object $c \in C$ there
exists a 1-simplex $[c \to t]$ in $\nerve(C)$.
\end{theorem}

\begin{proof}
The unique morphism $c \to t$ (from terminality) gives a
\texttt{ComposableArrows C 1}:
\begin{lstlisting}
theorem nerve_connected_to_terminal (t : C) (ht : IsTerminal t) (c : C) :
    exists e : ComposableArrows C 1, e.obj 0 = c /\ e.obj 1 = t :=
  <ComposableArrows.mk1 (ht.from c), rfl, rfl>
\end{lstlisting}
\end{proof}

\begin{theorem}[Edge uniqueness]\label{thm:unique}
Any two 1-simplices from $c$ to $t$ in $\nerve(C)$ are equal.
\end{theorem}

\begin{proof}
Since $t$ is terminal, any two morphisms $f, g : c \to t$ satisfy $f = g$:
\begin{lstlisting}
theorem nerve_edge_to_terminal_unique (t : C) (ht : IsTerminal t) (c : C) :
    forall (f g : c --> t), ComposableArrows.mk1 f = ComposableArrows.mk1 g := by
  intros f g; congr 1; exact ht.hom_ext f g
\end{lstlisting}
\end{proof}

\begin{theorem}[Retraction]\label{thm:retraction}
The nerve of $C$ admits a retraction pair with $\Delta[0]$: a projection
$\nerve(C) \to \Delta[0]$ and a section $\Delta[0] \to \nerve(C)$ picking the
terminal vertex.
\end{theorem}

Together, Theorems~\ref{thm:connected}--\ref{thm:retraction} establish that
the nerve of a category with a terminal object has connected and
simply-connected homotopy category---the categorical form of contractibility.
For nerves of categories (which are 2-coskeletal by a theorem in
Mathlib~\cite{Mathlib}), this is equivalent to topological contractibility of
the geometric realization.

\section{Relation to the broader architecture}\label{sec:broader}

The Quillen formalization is the \emph{topology discharge}; the algebraic lifting story is
developed primarily in~\cite{SpivackFiber}. We record two companion connections for context.

\paragraph{Embedding problems.}
An embedding problem in Galois theory asks whether a surjection
$H \twoheadrightarrow G$ admits a homomorphic section. This is exactly a
splitting question for the associated group extension
$1 \to \ker \to H \to G \to 1$. See~\cite{SpivackFiber}:

\begin{theorem}[Embedding problems as splitting]\label{thm:embed}
An embedding problem is solvable if and only if the associated group extension
splits, if and only if there exists a section with trivial cocycle.
\end{theorem}

\paragraph{Descent infrastructure.}
For an abelian group $A$, the doubling map $[2] : a \mapsto a + a$ has fibers
that are cosets of the 2-torsion subgroup $A[2]$. We prove that $A[2]$ acts
freely and transitively on each fiber (torsor structure)---the algebraic
foundation for 2-descent on elliptic curves. Separately, faithfully flat
descent provides a \emph{conservative} instance of the architecture: base
change reflects injectivity, surjectivity, and bijectivity, contrasting with
group extensions where the forgetful map is surjective but not conservative.

\section{Lean formalization details}\label{sec:lean}

The formalization consists of 21 Lean files in
\texttt{InfinityCompression/GeneralMethod/}, organized in four subdirectories:

\begin{center}
\footnotesize
\begin{tabular}{@{}llp{0.55\textwidth}@{}}
\toprule
\textbf{Directory} & \textbf{Files} & \textbf{Content} \\
\midrule
\texttt{Quillen/} & 7 &
  Fiber lemma, Galois connection case, order complex, nerve contractibility,
  prism operators, homotopy verification, \textbf{Theorem~A} \\
\texttt{GroupExtension/} & 10 &
  Fibers, cocycles, splitting criterion, local-global, central extensions,
  non-abelian obstruction, Mathlib bridge \\
\texttt{Descent/} & 2 &
  Faithfully flat descent, elliptic curve 2-torsion torsor \\
\texttt{Galois/} & 1 &
  Embedding problems as extension splitting \\
\bottomrule
\end{tabular}
\end{center}

All files compile with zero \texttt{sorry}. The formalization builds on
Mathlib's \texttt{GaloisConnection}, \texttt{nerve}, \texttt{ComposableArrows},
\texttt{IsTerminal}, \texttt{GroupExtension}, and \texttt{twoTorsionPolynomial}
APIs.

\section{Conclusion}\label{sec:conclusion}

We have presented a machine-checked formalization of the homotopy equivalence
data for Quillen's Theorem~A in the Galois connection case, together with
nerve contractibility for terminal categories, embedding problems as extension
splitting, and descent infrastructure. To our knowledge, this is the first
formalization of any version of Quillen's Theorem~A in a proof assistant, and
the first formalization connecting embedding problems to group extension
splitting with cocycle obstructions.

The suite needs this discharge because the architecture is not confined to algebraic
liftability: contractible fibers and nerve homotopy data show the same comparison pattern
organizing homotopy-theoretic equivalence statements.

For the Lean ecosystem specifically, the formalization contributes new
definitions (\texttt{QuillenAData}, \texttt{EmbeddingProblem},
\texttt{DoublingFiber}, \texttt{TwoTorsionSet}) and theorems
(\texttt{quillenA}, \texttt{solvable\_iff\_trivial\_cocycle},
\texttt{nerve\_connected\_to\_terminal}) that build on and extend Mathlib's
existing APIs for Galois connections, nerves, group extensions, and elliptic
curves.

The main open directions are: (1)~upgrading the edge-level homotopy data to a
full simplicial homotopy once Mathlib defines that notion, (2)~the general
Quillen Theorem~A for arbitrary contractible fibers, (3)~the full
$H^2 \leftrightarrow \text{extensions}$ bijection, and (4)~the bridge from
abstract embedding problems to actual Galois extensions via Mathlib's
\texttt{IntermediateField}.

\begin{thebibliography}{99}

\bibitem{Quillen1978}
D.~Quillen.
\newblock Homotopy properties of the poset of nontrivial $p$-subgroups of a
  group.
\newblock \emph{Advances in Mathematics}, 28(2):101--128, 1978.

\bibitem{BWW2004}
A.~Bj\"orner, M.~Wachs, and V.~Welker.
\newblock Poset fiber theorems.
\newblock \emph{Transactions of the American Mathematical Society},
  357(5):1877--1899, 2005.

\bibitem{Barmak2010}
J.~Barmak and E.~Minian.
\newblock On Quillen's Theorem~A for posets.
\newblock \emph{Journal of Combinatorial Theory, Series A}, 118(8):2445--2453,
  2011.
\newblock arXiv:1005.0538.

\bibitem{Lean4}
L.~de Moura and S.~Ullrich.
\newblock The Lean 4 theorem prover and programming language.
\newblock In \emph{CADE 28}, volume 12699 of \emph{LNCS}, pages 625--635.
  Springer, 2021.

\bibitem{Mathlib}
The Mathlib community.
\newblock The Lean mathematical library.
\newblock In \emph{CPP 2020}, pages 367--381, 2020.
\newblock \url{https://github.com/leanprover-community/mathlib4}.

\bibitem{SpivackFiber}
N.~Spivack.
\newblock Fiber architecture for group extensions in Lean~4: cocycles,
  splitting, and the cohomological bridge.
\newblock General-method manuscript, 2026.

\bibitem{SpivackSynthesis}
N.~Spivack.
\newblock \emph{Certification, Realization, and Obstruction: A Universal Fiber Architecture}.
\newblock Synthesis manuscript, 2026.

\bibitem{SpivackSummit}
N.~Spivack.
\newblock \emph{Reflective Non-Exhaustion: Certification, Realization, Fibers, and Obstruction}.
\newblock Summit manuscript, 2026.

\end{thebibliography}

\end{document}
