% Program synthesis paper — map of the Infinity Compression paper suite and Lean artifact.
\documentclass[11pt]{article}
\input{../suite_preamble}
\usepackage{graphicx}
\input{../suite_macros}

\title{Certification, Realization, and Obstruction:\\[0.35em]
\large A Universal Fiber Architecture}
\author{Nova Spivack}
\date{2026}

\begin{document}

\maketitle

\begin{abstract}
This paper closes the editorial loop for a multi-manuscript program in machine-checked
mathematics. The executive claim is that a single \emph{collapse / refinement / fiber}
architecture---certification at a bare carrier, strict comparison upstairs, obstruction
organized as fiber and section data---is instantiated internally in dependent type theory,
validated on twelve external mathematical families, and then discharged into algebra,
arithmetic (embedding problems), topology (Quillen for Galois connections), and
logic/computability (self-certification and halting anchors). We state the role of each
companion paper, a cross-domain dictionary, route labels used in the program (C/A/B/D),
and what remains open. The intended reader is an advanced mathematician who has not read
the series; citations point to the formal manuscripts and the Lean repository.
\end{abstract}

\SuiteSeriesContext{This manuscript is the \emph{synthesis map}.}

\keywords{formal verification, certification vs.\ realization, fibers, obstruction,
  Infinity Compression, Lean~4, program architecture}

\tableofcontents
\newpage

\section{Executive statement}

\textbf{One-sentence thesis.} Canonical certification at a collapsed bare carrier does not
exhaust structurally significant information: the gap is organized by a forgetful map,
nontrivial fibers, and explicit obstructions---and that pattern repeats across internal
formalization, external benchmarks, and classical mathematical domains.

\paragraph{What compresses.}
Quotient language, section language, obstruction language, and the implementation-vs.-certificate
distinction are often treated as separate methodological topics. This program shows they are
instances of one \emph{comparison architecture}: fix carriers, fix a forgetful map $\pi$,
ask where collapse occurs, and analyze fibers and obstructions when it does not. That
compression---from scattered idioms to one reusable pattern---is the synthesis-level
payoff.

\section{Certification, realization, and non-exhaustion}

\paragraph{Certification.}
\emph{Certification} means a theorem-level invariant packaged as a canonical record at a
\emph{bare} carrier (proof-irrelevant at that carrier when the type forces it): many
derivation routes may exist, but the official statement names a single certified outcome.
It does \emph{not} mean informal ``approval'' or social certification; it means the
library's designated bare certificate type and its collapse theorems.

\paragraph{Realization.}
\emph{Realization} means structured or proof-relevant data above that bare layer: witnesses,
roles, sections, splittings, nerve data, or fiber points that remain mathematically
load-bearing once $\pi : E \to B$ is fixed. It is not mere ``extra syntax'': it is the
carrier on which non-injectivity and fiber lemmas become meaningful.

\paragraph{Non-exhaustion.}
\emph{Non-exhaustion} is the proved claim that certification does not determine realization:
distinct legitimate data can share the same bare image under $\pi$, organized by fibers and
obstruction laws. It does \emph{not} assert that every formal system exhibits this
phenomenon; it is established on \emph{named} carriers and explicit hypotheses in the
artifact.

\paragraph{Examples (suite instances).}
Flagship Infinity Compression: collapse at the bare summit record; non-injectivity of
$\pi_A$ on enriched splits~\cite{SpivackIC}. Group extensions: splitting iff trivial
cocycle~\cite{SpivackFiber}. Benchmark families: structured bundles beyond flat
baselines~\cite{SpivackValidation}. Quillen / Galois connections: contractible fibers
feeding Theorem~A data~\cite{SpivackQuillen}.

\section{Flagship internal theorem}

The Infinity Compression flagship paper~\cite{SpivackIC} is the \emph{internal origin
theorem}: collapse of standard extraction, reflective split, strict refinement of the
typed forgetful map, and partial fiber classification---all in the library's native
carriers.

\section{External validation (transferability)}

The twelve-tranche survey~\cite{SpivackValidation} is the \emph{transferability paper}: the
same schema applies to quotients, products, dependent sums, coproducts, subtypes,
orbit and ideal quotients, pushouts, classifiers, localization, pullbacks, and a
non-surjective control. Its job is not theorem-level leverage in classical fields; it
rules out idiosyncrasy.

\section{Algebraic and arithmetic engine}

The group-extension paper~\cite{SpivackFiber} is the \emph{algebraic and arithmetic engine}:
section cocycles, splitting iff trivial cocycle, embedding problems as extension splitting,
local-to-global obstruction, and the documented bridge toward Galois-theoretic towers.

\section{Topological discharge}

Quillen's Theorem~A for Galois connections~\cite{SpivackQuillen} is the \emph{topology
discharge}: nerves, contractible fibers over down-sets, and homotopy-equivalence data at
the simplicial level for the Galois-connection case.

\section{Logic and computability anchor}

Route~D in the program supplies a computability-facing discharge: self-certification
phenomena aligned with standard undecidability infrastructure (e.g.\ halting and Rice
scaffolding in the Mathlib \texttt{ComputablePred} layer), making the analogy to
classical incompleteness precise without identifying the main theorem with G\"odel's
syntax. Exact statements live in the Lean \texttt{RouteD/} and \texttt{Summit/} modules;
this synthesis paper does not restate every definition.

\section{Cross-domain dictionary}

The following table is a navigational index, not a formal commuting diagram.

\begin{center}
\resizebox{\textwidth}{!}{%
\footnotesize
\begin{tabular}{@{}p{1.1cm}p{1.7cm}p{1.7cm}p{1.9cm}p{2.2cm}@{}}
\toprule
\textbf{Domain} & \textbf{Bare / cert.} & \textbf{Enriched} & \textbf{Comparison} & \textbf{Headline / obstruction} \\
\midrule
IC flagship & canonical bare cert. & enriched split & $\pi_A$ & non-inj.; nontrivial fiber \\
T1--T12 & flat baseline & structured bundle & projection & fibers, sections (T12: boundary) \\
Group ext. & section exists & splitting data & section data & split $\Leftrightarrow$ trivial cocycle \\
Embedding & quotient data & liftings & extension cmp. & solvable $\Leftrightarrow$ trivial cocycle \\
Quillen GC & nerve image & homotopy witn. & nerve maps & contractible fibers \\
Computability & cert.\ layer & realized behav. & cert.\ cmp. & halting / Rice \\
\bottomrule
\end{tabular}%
}
\end{center}

\section{Routes C, A, B, and D (program labels)}

\begin{itemize}[leftmargin=2em]
  \item \textbf{Route C (headline):} reflective non-exhaustion at the flagship carriers.
  \item \textbf{Route A (infrastructure):} supporting general lemmas (e.g.\ collision
    patterns, forgetful infrastructure) used by the headline.
  \item \textbf{Route B (applications):} algebra, topology, arithmetic bridges---the
    general-method papers.
  \item \textbf{Route D (logic/computability):} halting and Rice-style scaffolding aligned
    with computability theory rather than informal metaphor.
\end{itemize}

\section{Why this is a reusable research instrument}

The architecture is portable because it isolates \emph{where} collapse occurs, \emph{where}
comparison must move upstairs, and \emph{what} invariant obstructs lifting---independent
of any single application domain. Lean provides a checkable artifact; the paper suite
separates internal origin, transferability, domain discharges, and summit-level claims.

\paragraph{Operational recipe.}
\begin{enumerate}[label=\arabic*., leftmargin=2em]
  \item Identify the bare certificate carrier $B$.
  \item Identify an enriched carrier $E$ tracking realization data.
  \item Define the forgetful or comparison map $\pi : E \to B$.
  \item Prove or locate \emph{collapse}: when many routes hit one bare point in $B$.
  \item Test injectivity of $\pi$; if it fails, work on fibers $\pi^{-1}(b)$.
  \item Extract obstruction laws (cocycle, contractibility, non-certifiability) governing
    liftability and equivalence.
\end{enumerate}

\section{What this architecture is not}

\begin{itemize}[leftmargin=2em]
  \item \textbf{Not} a claim that all of mathematics reduces to one master theorem.
  \item \textbf{Not} a generic restatement of dependent-type proof relevance: the
    contribution is \emph{where} collapse and non-injectivity are proved, on named carriers.
  \item \textbf{Not} mere packaging rhetoric: the suite contains formal theorems with
    zero \texttt{sorry} in the scoped modules.
  \item \textbf{Not} (yet) a single universal quantified Lean theorem over all domains at
    once: that unification remains a frontier goal.
\end{itemize}

\section{Portability vs.\ inevitability}

The current suite establishes strong \emph{portability}: the same comparison pattern
reappears across internal, benchmark, and classical discharges. A stronger
\emph{inevitability} claim---that every reflective certification setup must exhibit
non-exhaustion in a single abstract class---remains a summit-level ambition, not a
finished theorem.

\section{What remains open}

Full fiber classification in the flagship setting; complete simplicial homotopy
formalization behind Quillen; full $H^1$/$H^2$ classification in Mathlib; stronger
unified quantification across all carriers in a single declaration; publication and Mathlib
submission scheduling.

\section*{Acknowledgments}

Repository: \url{https://github.com/novaspivack/infinity-compression-lean}.

\bibliographystyle{alpha}
\begin{thebibliography}{99}

\bibitem{SpivackIC}
N.~Spivack.
\newblock \emph{Canonical Certification Does Not Exhaust Reflective Structure:
  Reflective Split, Strict Refinement, and Fiber Structure in Infinity Compression}.
\newblock Flagship manuscript, 2026.

\bibitem{SpivackValidation}
N.~Spivack.
\newblock \emph{External Validation of a Positive-Closure Proof Architecture:
  A Multi-Tranche Survey}.
\newblock Validation manuscript, 2026.

\bibitem{SpivackFiber}
N.~Spivack.
\newblock \emph{Fiber Architecture for Group Extensions in Lean~4: Cocycles, Splitting,
  and the Cohomological Bridge}.
\newblock General-method manuscript, 2026.

\bibitem{SpivackQuillen}
N.~Spivack.
\newblock \emph{Quillen's Theorem~A for Galois Connections: A Machine-Checked
  Formalization in Lean~4}.
\newblock General-method manuscript, 2026.

\bibitem{SpivackSummit}
N.~Spivack.
\newblock \emph{Reflective Non-Exhaustion: Certification, Realization, Fibers, and Obstruction}.
\newblock Summit manuscript, 2026.

\end{thebibliography}

\end{document}
